A Wide-Coverage Symbolic Natural Language Inference System

### 2019-03-04 paper2018/paper.tex
change paper to times; add back some text

\documentclass[11pt]{article}
\usepackage{acl2019}
\usepackage{todonotes}
\usepackage{tabularx}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

@@

\paragraph{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP → CN → CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete syntax of various natural languages.
% These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity.

@@

All the underlying systems (GF, Haskell, Coq) are based on lambda
calculi with types. We take advantage of typing, ensuring that each
translation preserve typing, locally:
\begin{enumerate}[noitemsep]
\item Every GF syntactic category $C$ is mapped to a type noted $⟦C⟧$.
\item GF Functional types are mapped compositionally : $⟦A → B⟧ = ⟦A⟧ → ⟦B⟧$
\item Every GF syntactic construction function ($f : X$) is mapped to a function $⟦f⟧$ such that $⟦f⟧ : ⟦X⟧$.
\item GF function applications are mapped compositionally: $⟦t(u)⟧ = ⟦t⟧ (⟦u⟧)$.
\end{enumerate}
Because all systems embed the simply-typed lambda calculus, ensuring
type-preservation locally means that types are preserved globally.
Therefore, we are certain that every GF syntax tree can be mapped to
Haskell, and eventually Coq, without error.

@@

\paragraph{definites} A naive way to handle definites is using an
existential type. However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$∀x. person(x) → ∃y. dog(y) ∧ pet(x,y)$.
%
Instead, our take is that definites should be treated as an anaphoric
expression with an implicit referrent. That is, if the referent is not
found in the discourse, then it will be forcibly introduced, using an
existential type, \emph{at the top-level} of the expression. To be
able to do this, we record all definites without referent, using
another portion of the environment (using a monadic effect).
%SC
%For the
%above example, we obtain the desired interpretation:
%$∃y. dog(y) ∧ (∀x. person(x) → pet(x,y))$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost θ orders. θ $\ge$ 0.  ITEL won at least θ+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.
%SC
%Finally, ``more'' is intepreted as ``at least
%<elliptic quantity>+1''. This treatment, which we illustrated here on
%an example, is systematic in our implementation.

\paragraph{Adjectives}
We interpret gradable adjectives using a pair of a
measure $m : objects → Z$ and a threshold $τ : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $τ$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$∀x. m_1 x = -m_2 x$ and thresholds do not overlap $τ_1 + τ_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Kennedy:2007}.  Additionally
adjectival predicates, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable.
% SC
%Indeed, the tactic that Coq offers for this purpose can
%solve all such problems in the FraCaS suite, automatically.

% \paragraph{Adverbs}
% Another point, of minor theoretical importance but major practical
% one, is our handling of adverbial phrases. We interpret adverbs (and
% in general all adverbial and prepositional phrases) as VP-modifiers:
% $Adv = VP → VP$, where $VP = object → Prop$. However, applying adverbs
% to verb-phrases heavily complicates the Coq proofs, because such
% phrases can contain quantifiers. Therefore, we instead move the
% adverbs, so that they apply to (atomic) verbs only. Proofs can then be
% simplify accordingly. \todo{theoretical justification in the
%   literature?}

@@

% \paragraph{error analysis}
% Our system fails to classify correctly 28 cases out of 259.  We give
% here a summary of the missing features which are responsible for the
% biggest number of failures. The one biggest source of error is an
% incomplete handling of group readings. (FraCaS
% 013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
% are cases where a syntactic conjunction of individuals is treated as a
% semantic group, or where precise counting of the members of a group is
% necessary.  Other problematic cases are where a definite plural does
% not have a universal reading (091, 094, 095) --- these require an
% analysis at the pragmatic level, which our system lacks. Additionally,
% neither measure phrases (242) nor attributive comparatives (244, 245)
% are handled.

>>>>

\documentclass[11pt]{article}
\usepackage{acl2019}
\usepackage{times}
\usepackage{todonotes}
\usepackage{tabularx}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

@@

\paragraph{GF}
In GF, abstract syntax is comprised of: a) a number of syntactic
categories, and b) a number of syntactic construction functions. The latter
provide the means to compose basic syntactic categories into more complex
ones.  For example, consider the constructor: $AdjCN : AP \to CN \to CN$. This  expresses
that one can append an adjectival phrase to a common noun and obtain
a new common noun. Furthermore,  GF is equipped  with a library of mappings from abstract
syntax to the concrete syntax of various natural languages.
% These mappings can be inverted by GF, thus offering parsing from natural text into abstract syntax.
However, in this project we skip the
parsing phase and use the parse trees constructed by \citet{Ljunglof:2012},
thereby avoiding any syntactic ambiguity.

@@

All the underlying systems (GF, Haskell, Coq) are based on lambda
calculi with types. We take advantage of typing, ensuring that each
translation preserve typing, locally:
\begin{enumerate}[noitemsep]
\item Every GF syntactic category $C$ is mapped to a type noted $\llbracket{}C\rrbracket{}$.
\item GF Functional types are mapped compositionally : $\llbracket{}A \to B\rrbracket{} = \llbracket{}A\rrbracket{} \to \llbracket{}B\rrbracket{}$
\item Every GF syntactic construction function ($f : X$) is mapped to a function $\llbracket{}f\rrbracket{}$ such that $\llbracket{}f\rrbracket{} : \llbracket{}X\rrbracket{}$.
\item GF function applications are mapped compositionally: $\llbracket{}t(u)\rrbracket{} = \llbracket{}t\rrbracket{} (\llbracket{}u\rrbracket{})$.
\end{enumerate}
Because all systems embed the simply-typed lambda calculus, ensuring
type-preservation locally means that types are preserved globally.
Therefore, we are certain that every GF syntax tree can be mapped to
Haskell, and eventually Coq, without error.

@@

\paragraph{definites} A naive way to handle definites is using an
existential type. However, if the semantics does not feature a
dynamic element, then the existential quantification is introduced
locally. This means that the quantifier can be introduced in the wrong
context. Consider the phrase ``everyone pets the dog''. The structure
of the interpretation would be
$\forall{} x. person(x) \to \exists{}y. dog(y) \wedge pet(x,y)$.
%
Instead, our take is that definites should be treated as an anaphoric
expression with an implicit referrent. That is, if the referent is not
found in the discourse, then it will be forcibly introduced, using an
existential type, \emph{at the top-level} of the expression. To be
able to do this, we record all definites without referent, using
another portion of the environment (using a monadic effect).
For the
above example, we obtain the desired interpretation:
$\exists{}y. dog(y) \wedge (\forall{}x. person(x) \to pet(x,y))$.

\paragraph{Phrasal comparatives}
Previous attempts to tackle the section of the FraCaS test suite
devoted to comparatives showed that handling them is not easy. Our
strategy here is to leverage our dynamic semantics, revealing an
anaphoric element of comparatives. Indeed, Consider the hypothesis of
(FraCaS 239): ``ITEL won more orders than APCOM lost.''  We postulate
that this sentence is equivalent to the following two separate parts:
``APCOM lost zero or more orders. ITEL won more orders [than some
elliptic quantity].''  A quantity is introduced every time we talk
about some quantity (indexed by a CN, in this case ``orders''), and it
can be referred to by a comparative, later in the discourse. Using
this idea, we can go one level deeper in the interpretation of our
example: ``APCOM lost $\theta$ orders. $\theta$ $\ge$ 0.  ITEL won at least $\theta$+1
orders.''. We see here how the quantities are introduced. They are
added to the environment so that, they can be referred to as elliptic
quantity expressions.
% SC
%Finally, ``more'' is systematically intepreted as ``at least
%<elliptic quantity>+1''. This treatment, which we illustrated here on
%an example, is systematic in our implementation.

\paragraph{Adjectives}
We interpret gradable adjectives using a pair of a
measure $m : objects \to Z$ and a threshold $\tau : Z$, where $Z$ is
treated as an abstract ordered ring by Coq. (This structure has no
dynamic aspect in our model, and thus is entirely handled within Coq.)
For subsective adjectives, $\tau$ will additionally depend on the class
of the object in question. This structure has the benefit that
opposite adjectives can be easily represented (measures are opposites
$\forall{}x. m_1 x = -m_2 x$ and thresholds do not overlap $\tau_1 + \tau_2 >
0$). Formalization aside, this idea is reminiscent of degree-based
approaches to gradable adjectives of
\citet{Cresswell:1976,Kennedy:2007}.  Additionally
adjectival predicates, as present in the FraCaS suite, are interpreted as
linear inequations in $Z$. Solving systems of such inequations is
decidable.
Indeed, the tactic that Coq offers for this purpose can
solve all such problems in the FraCaS suite, automatically.

% \paragraph{Adverbs}
% Another point, of minor theoretical importance but major practical
% one, is our handling of adverbial phrases. We interpret adverbs (and
% in general all adverbial and prepositional phrases) as VP-modifiers:
% $Adv = VP \to VP$, where $VP = object \to Prop$. However, applying adverbs
% to verb-phrases heavily complicates the Coq proofs, because such
% phrases can contain quantifiers. Therefore, we instead move the
% adverbs, so that they apply to (atomic) verbs only. Proofs can then be
% simplify accordingly. \todo{theoretical justification in the
%   literature?}

@@

\paragraph{error analysis}
Our system fails to classify correctly 28 cases out of 259.  We give
here a summary of the missing features which are responsible for the
biggest number of failures. The one biggest source of error is an
incomplete handling of group readings. (FraCaS
013, 014, 046, 084, 111, 124, 126, 127, 137, 171, 172, 191, 193, 195, 243, 250, 333, 346). These
are cases where a syntactic conjunction of individuals is treated as a
semantic group, or where precise counting of the members of a group is
necessary.  Other problematic cases are where a definite plural does
not have a universal reading (091, 094, 095) --- these require an
analysis at the pragmatic level, which our system lacks. Additionally,
neither measure phrases (242) nor attributive comparatives (244, 245)
are handled.
